\topic{14. Collision Detection}

\name{Erwin Coumans}: creater of Tiny Differentiable Simulator 
which is a header-only C++ and CUDA physics library 
for reinforcement learning and robotics with zero dependencies.

\subsubtitle{Classes of Objects \& Problems}
\lis{
  \item 2D vs. 3D
  \item Convex (random 2 point connect will not go out of boundary) vs. Non-Convex
  \item Polygonal (triangle, quads) vs. Non-Polygonal
  \item Open surfaces vs. Closed volumes 
  \item Geometric (thin-shell) vs. Volumetric
  \item Rigid vs. Non-rigid (deformable/flexible)
  \item Pariwise vs. Multiple (N-Body)
  \item CSG (Constructive solid geometry: intersection, union, complement) vs. B-Rep (polygonal mesh: list of vertices and triangles)
  \item Static vs. Dynamic
  \item And so on ... This may include other geometric representation schemata, etc.
}

\subtitle{Bounding Volumes (BVs)}
\lis{
  \item Sphere, AABB, OBB, 6-dop, Convex Hull
  \item $\rightarrow$ incresing complexity \& tightness of fit
  \item $\leftarrow$ decresing cost of (overlap tests + BV update)
}

\subtitle{Bounding Volumes Hierarchies} \\
\subsubtitle{Model Hierarchy:}
\lis{
  \item each node has a simple volume that bounds a set of triangles
  \item children contain volumes that each bound a different portion of the parent's triangles
  \item The leaves of the hierarchy ususally contain indicidual triangles (or the number of triangles in the leaf below a threshold)
}
\subsubtitle{Bounding Volume Test Tree (BVTT):}
\lis{
  \item recurse on both tree
  \item recurse on higher level tree (better but why?)
}
\subsubtitle{Principal Component Analysis (PCA):}
\lis{
  \item Algorithm to split
  \item Use to find the normal vector of an Objects
  \item data: $x_i, i=1,...,N$
  \item center: $c = \frac{1}{N}\sum_{i=1}^{N}x_i$
  \item $y_i = x_i - c, i = 1, ..., N$
  \item Covariance Matrix: $A=\frac{1}{N}\sum_{i=1}^{N}y_iy_i^{T}$ 
  \item $A^T=A$ is symmetric matrix because $(y_iy_i^T)T=(y_i^T)^T\cdot y_i^T=y_i \cdot y_i^T$
  \item eigendecomposition $A=U\Lambda U^T$ where $A$ is 3*3 symmetric, $U$ is 3*3 Orthogonal, $\Lambda$ is 3*3 diagonal
  \item \( \Lambda = 
  \begin{bmatrix}
  \lambda_1 &  &  \\
   & \lambda_2 &  \\
   &  & \lambda_3
  \end{bmatrix}
  \) where $\lambda_1 \geq \lambda_2 \geq \lambda_3 \geq 0$
  \item $U = [D * *]$ where $D$ is desired direction
}
\subsubtitle{Self Collision Detection:}
\lis{
  \item Copy the BVH then compare
  \item Another way is to brute force (compare every pair of triangles except neighboring ones) and parallel computing
}

\subtitle{Spatial Data Structures \& Subdivision}
\lis{
  \item Uniform Spatial Sub: uniformly split among the space
  \item Quadtree (2D) / Octree (3D): split evenly into 4, and concentrate into one grid
  \item kd-tree: similar to quad tree but split unevenly, and only split in 1D, common in AI
  \item BSP-tree: kd-tree but accept tilted split
}
\subsubtitle{Uniform Spatial Subdivision}
\lis{
  \item Decompose the objects (the entire simulated environment) into identical cells arranged in a fixed, regular grids (equal size boxes or voxels)
  \item To represnet an object, only need to decide which cells are occupied. To perform collision detection, check if any cell is occupied by two objects
  \item Storage: to represent an object at resolution of n voxels per dimension requires upto $n^3$ cells
  \item Accuracy: solids can only be "approximated"
  \item step 0 : clear the list
  \item step 1 : go over the first object, add element to grid's list
  \item step 2 : do the samething for the second object
  \item step 3 : for every cell, if two lists are non-empty, pair-wise collision checking
  \item self collision detection: for every cell, if the list has 2+ non-neighboring entries, pair-wise collision checking
  \item issue: space memory (need to store empty cell) and speed (need to go over all cells)
  \item solution: hash table with psudo-randomness
  \item step 1 : convert every grid to a very long table (e.g. index $=$ row $*$ num horizontal cell $+$ collumn)
  \item step 2 : condense it to a small table using a hash function
  \item step 3 : erase the long table
  \item for each 
}
\subtitle{signed distance field}
\lis{
  \item \name{Six-DoF Haptic Rendering of Contact between Geometrically Complex Reduced Deformable Models}
  \item for each grid, store closest distance $A$ to the object
  \item + outside, - inside
  \item for one point, if no signed distance field, check if $A \cdot N > 0 \rightarrow$ outside, or $A$ and $N$ has the same direction
  \item if signed distance field, read 8 points from memory
  \item bilinear in 2D, trilinear in 3D 
}
\subsubtitle{Contact force:} \equ{F_C = -k_CdN} 
\lis{
  \item $-k_C$ penalty stiffness, $d$ signed distance field value ($d < 0$ when in contact)
  \item Summ over all points
  \item Torque = $r \times F_C$
  \item Also compute gradients
}

\subsubtitle{Distributed Contact}
\lis{
  \item difficult contact - multiply contact point
  \item one is point shell (deformable), one is signed distance (rigid)
}
