\topic{Other Topics}

\subtitle{Digital Twin}
\lis{
  \item Digital twin: a digital copy of a human/animal/buiding...
  \item Digital humans: input an image and output a digital human
  \item \name{Apic games - Metahuman}: have a library of data and artist create a digital human based on those
  \item pro: easier to communicate in long distance
  \item con: security, law
}

\subtitle{Metaverse} \chinese{元宇宙}
\lis{
  \item differ from Multiverse \chinese{多元宇宙}
  \item Interact with each other in VR
  \item "2D Social Media" $\rightarrow$ Metaverse: the next generation Facebook, Zoom, Apple Facetime, Microsoft Teams, Google Meet etc.
  \item Cybersecurity, Global Warming, Digital Twins
  \item Pro: 
  \item Go on dates with long distance partners/spouses in VR
  \item Meet friends and family in VR
	\item Business meeting in VR $\rightarrow$ less airplane travel -> less CO2 emitted -> slow down global warming
	\item Doctors meeting patients in far away places in VR
  \item Shop in VR, attend concerts, school in VR, etc.
	\item Kardashev Type 1 civilization "world communication system": 
  0: Nothing, 0.5: We are here, 1: Control the planet, 2: Control the solar system, 3: Control the galaxy
}

%\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
% License
%\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
\subtitle{License}
\lis{
  \item MIT, BSD: similar to PDL, except they cannot use you to advertices (prefered)
  \item LGPL: library 
  \item GPL: if take code from GPL, then have to license under GPL. (company do not want this)
  \item Public Domain Licence: everyone can take it and I don't care (abuse)
}

%\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
% Important People
%\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\
\topic{Important People}

\subtitle{Lisa Su}
\lis{
  \item CEO of AMD
  \item Age: 55
  \item Country of birth: Taiwan
  \item Undergraduate, MS and PhD degrees at MIT
  \item Oct 2014 - Feb 2025: 76x AMD stock growth (57\% per year!)
  \item "Run toward the hardest problems. This approach has helped me to learn a tremendous amount from both success and failure."
}

\subtitle{Edwin Catmull}
\lis{
  \item Co-Founder, Pixar
  \item Preident, Walt Disney Animation Studios
  \item Turning Awars, 2019
  \item Age: 78
  \item Country of birth: United States
  \item Undergrauate, PhD at Univ. of Utah (Physics and CS)
  \item Invented texture mapping, (co-invented) z-Buffering, Catmull-Rom spline
  \item "Failure isn't a necessary evil. In fact, it isn't evil at all. It is a necessary consequence of doing something new."
  \item "If you give a good idea to a mediocre team, they will screw it up. If you give a mediocre idea to a brilliant team, they will either fix it or throw it away and come up with something better."
  \item "You are not your idea, and if you identify too closely with your ideas, you will take offense when ehty are challenged."
}

\subtitle{Hayao Miyazaki} \chinese{宫崎骏}
\lis{
  \item Co-Founder, Director, Studio Ghibli: \name{My Neighbor Totoro, Proco Posso, Spirited Away, ...}
  \item Animator, film-maker, manga artist
  \item "Godfather of animation in Japan" 
  \item Super famous. Numerous awards for his films
  \item Country of birth: Japan. Age: 84 
  \item "I strongly feel that AI is an insult to life itself."
  \item "Do everything by hand, even when using the computer."
  \item "I would like to make a film to tell children: 'It's good to be alive.' "
  \item "Violence is innate in humans. The real issue, is how to control it."
  \item "No cuts." Famously objected to Harvey Weinstein cutting his films.
  \item "I think 2D animation disappeared from Disney because they made so many uninteresting films. It's too bad. I thought 2D and 3D could coexist happily."
}


